Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-06-04 Manuscript/manuscript.Rmd
manuscript edits

Using the sets of cell line specific genomic annotations, we created three types of predictors that were used to describe the spatial relationship with genomic bins, $X = \{X_1, X_2, ..., X_p\}$, where $p$ was the complete set of functional annotations. They are described in detail below:

@@

The cell line- and resolution-specific data, composed of the responsed vector and predictor-type specific data matrix  then split into a 7:3 ratio of training to testing sets. Furthermore, for the task of performance evaluation and prediction, the test set remained constant for all ensemble systems to account for an unbiased comparison.

@@

In order to suitably reduce the feature space while also maintaining the inclusion of possibly colocalized annotations, we implemented an elastic-net regularization technique prior to modelling. The technique using a combination of the *l_{1}* and *l_{2}* penalties. For each elastic-net regularization, 10-fold cross-validation was used to tune the parameters that controlled the penalty terms. <!--The parameters that maximized the mean AUPRC across each cross-fold was chosen.--> All elastic-net algorithms were implemented in R using the *caret* package version 6.0 [@kuhn2012caret]. 
-->

@@

Once the appropriate predictor type and set of features were determined, a final set of random forest classification models were built to compare model performances between re-sampling techniques and predictor types. <!--To reduce bias due to random dataset generation, 10-fold cross validation was used to tune the number of features to consider at each node and the number of ensemble trees to aggregate.--> The default number of features to consider at each node of the random forest algorithm was set as $\sqrt{p}$, where *p* refers to the number of predictors included in the model. Likewise, the number of ensemble trees to aggregate was set at 500. All random forest algorithms were implemented in R using the *caret* package version 6.0 [@kuhn2012caret].

The performance of each ensemble system in the framework was compared using the composite metric, Matthews Correlation Coefficient. The MCC is defined as follows:

@@

To establish the feature space for all downstream models, we constructed 3 different predictor-types. These predictor-type specific annotations represented the data matrix, $X = \{X_1, X_2, ..., X_p\}$, where $p$ was the complete set of functional annotations [?? $p$ is actually the total number of functional annotations, $X_i$ is the functional annotation itself ??]. Overlap counts was the least granular [?? Maybe define granular ??] predictor-type considered, simply defined by the number of annotation-specific regions located within each genomic bin (Figure 2A). Overlap percents was a more refined [?? what is refined in this context ??] predictor-type, defined as the total percent overlap between each annotation-specific region and genomic bins (Figure 2B). Lastly, distances, the most granular predictor-type, described the spatial relationship between the center of each genomic bin to the center of the nearest annotation-specific region (Figure 2C). A $log_{2}}$ transformation was used as a normalization technique in order to account for the skewness present in the distance densities (Supplementary Figures 1-2). [?? Why would skewness matter, why not use raw data ??]

>>>>

Using the sets of cell line specific genomic annotations, we created three types of predictors that were used to describe the spatial relationship with genomic bins, $\textbf{X} = \{X_1, X_2, ..., X_p\}$, where $p$ was the complete set of functional annotations. They are described in detail below:

@@

The cell line- and resolution-specific data, composed of the response vector and predictor-type specific data matrix (**X**), was then split into a 7:3 ratio of training to testing sets. For the task of performance evaluation and prediction, the test set remained constant to account for an unbiased comparison between data level characteristics.

@@

In order to suitably reduce the feature space while also maintaining the inclusion of possibly colocalized annotations, we implemented an elastic-net regularization technique prior to modelling. The technique using a combination of the *l_{1}* and *l_{2}* penalties. For each elastic-net regularization, 10-fold cross-validation was used to tune the parameters that controlled the penalty terms. All elastic-net algorithms were implemented in R using the *caret* package version 6.0 [@kuhn2012caret]. 
-->

@@

random forest classification models were built to compare model performances between re-sampling techniques and predictor types. To reduce bias due to random dataset generation, 10-fold cross validation was used to tune the number of annotations, *p*, to consider at each node. The number of trees that were aggregated for each random forest model was set to the default, n=500. All random forest algorithms were implemented in R using the *caret* package version 6.0 [@kuhn2012caret]. Once the model was implemented, it was validated on the test set. The performance of each ensemble system in the framework was compared using the composite metric, Matthews Correlation Coefficient (MCC), which is defined as follows:

@@

To establish the feature space for all downstream models, we constructed 3 different predictor-types. These predictor-type specific annotations represented the data matrix, $\textbf{X} = \{X_1, X_2, ..., X_p\}$, where $p$ was the complete set of functional annotations [?? $p$ is actually the total number of functional annotations, $X_i$ is the functional annotation itself ??]. Overlap counts was the least granular [?? Maybe define granular ??] predictor-type considered, simply defined by the number of annotation-specific regions located within each genomic bin (Figure 2A). Overlap percents was a more refined [?? what is refined in this context ??] predictor-type, defined as the total percent overlap between each annotation-specific region and genomic bins (Figure 2B). Lastly, distances, the most granular predictor-type, described the spatial relationship between the center of each genomic bin to the center of the nearest annotation-specific region (Figure 2C). A $log_{2}}$ transformation was used as a normalization technique in order to account for the skewness present in the distance densities (Supplementary Figures 1-2). [?? Why would skewness matter, why not use raw data ??]
